---
title: "Naive Bayes Classification Assignment"
format: html
editor: visual
---

## Assignment Description

This assignment is designed to test your knowledge of Naive Bayes Classification. It closely mirrors our [naive_bayes_penguins.qmd](https://github.com/NSF-ALL-SPICE-Alliance/DS400/blob/main/week7/naive_bayes_penguins.qmd) from lectures 10/1 and 10/3. We reflect back on the true vs fake news dataset from the beginning of the semester and apply the new skills in our bayesian toolbox.

This assignment is worth 16 points and is due by 10:00am on October 15th. Each section has a number of points noted. To turn in this assignment, render this qmd and save it as a pdf, it should look beautiful. If you do not want warning messages and other content in the rendered pdf, you can use `message = FALSE, warning = FALSE` at the top of each code chunk as it appears in the libraries code chunk below.

### Load Libraries

```{r, message=FALSE, warning=FALSE}
library(bayesrules)
library(tidyverse)
library(e1071)
library(janitor)
library(ggplot2)
```

### Read in data

```{r}
data(fake_news)
```

```{r}
fake_news
```

### Challenge

[**Exercise 14.7**](https://www.bayesrulesbook.com/chapter-14#exercises-13) **Fake news: three predictors**

Suppose a ***new news article*** is posted online – it has a 15-word title, 6% of its words have negative associations, and its title *doesn’t* have an exclamation point. We want to know if it is fake or real

### Visualization (Exploratory Data Analysis) - 2 points

Below, insert a code chunk(s) and use `ggplot` to visualize the features of the data we are interested in. This can be one or multiple visualizations

-   Type (fake vs real)

```{r}
ggplot(data = fake_news, aes(x = type, fill = type)) +
  geom_bar()
```

```{r}
ggplot(fake_news, aes(x = type)) +
  geom_bar(fill = c("pink", "orange"), alpha = 0.7) +
  geom_text(stat = 'count', aes(label = ..count..), vjust = -0.5) +
  labs(title = "Count of Real vs Fake", x = NULL, y = "Count") +
  theme_minimal()
```

-   Pie Chart : to give me another visualization of real vs. fake to better understand it

```{r}
ggplot(fake_news, aes(x = "", fill = type)) +
  geom_bar(width = 1) +
  coord_polar(theta = "y") +
  labs(title = "Real vs Fake") +
  theme_void() 
```

-   Number of words in the title (numeric value)

```{r}
ggplot(fake_news, aes(x = title_words)) +
  geom_histogram(binwidth = 1, fill = "purple", color = "red", alpha = 0.7) +
  labs(title = "Distribution of Words in Titles", x = "Words per Title", y = "Frequency") +
  theme_minimal()
```

-   Negative associations (numeric value)

```{r}
ggplot(fake_news, aes(x = negative, fill = type)) +
  geom_histogram(bins = 30, position = "dodge", alpha = 0.7) +
  labs(title = "Distribution of Negative Associations by News Type",
       x = "Negative Association Score",
       y = "Count") +
  scale_fill_manual(values = c("real" = "green", "fake" = "pink")) +
  theme_minimal()
```

-   Exclamation point in the title (true vs false)

```{r}
title_count <- nrow(fake_news)
cat("The number of titles in the fake_news dataset is:", title_count, "\n")
```

I just used this code to double check how many titles there are in the fake_news data set to see if the results to my visualization is correct and accurate.

-   Creating a new column indicating that there is ! points & without ! points

```{r}
fake_news <- fake_news %>%
  mutate(title_has_excl = str_detect(title, "!"))

exclamation_counts <- fake_news %>%
  count(title_has_excl, name = "count") 
```

### Bar Plot of Title with Exclamation Points

```{r}
ggplot(exclamation_counts, aes(x = factor(title_has_excl, labels = c("No", "Yes")), y = count)) +
  geom_bar(stat = "identity", fill = "yellow", alpha = 0.7, width = 0.6) +
  geom_text(aes(label = count), position = position_stack(vjust = 0.5), size = 5) +
  labs(title = "Count of Titles with Exclamation Points", x = "Has ! Point", y = "Count") +
  theme_minimal(base_size = 15)
```

Stacked Bar

### Interpretation of Visualization - 2 points

Below, write a few sentences explaining whether or not this ***new news article*** is true or fake solely using your visualization above

-   After looking at my visualizations that I did for the data set it is pretty clear that the news article is real. As the real always outweighs the fake in the variables that I used for my visualizations above.

### Perform Naive Bayes Classification - 3 points

Based on these three features (15-word title, 6% of its words have negative associations, and its title *doesn’t* have an exclamation point), utilize naive Bayes classification to calculate the posterior probability that the article is real. Do so using `naiveBayes()` with `predict()`.

Below, insert the code chunks and highlight your answer

### Break Down the Model - 5 points

Similar to the penguins example, we are going to break down the model we created above. To do this we need to find:

### Naives Bayes Classification

```{r}
naive_model_fake <- naiveBayes(type ~ title_words + title_has_excl + negative, data = fake_news)
```

Our_fake is the name of my new data_set

```{r}
our_fake <- data.frame(title_words = 15, negative = 0.06, title_has_excl = FALSE)
view(our_fake)
```

Ask model to predict

```{r}
predict(naive_model_fake, newdata = our_fake, type = "raw")
```

### Test my Model

```{r}
fake_news <- fake_news %>% 
  mutate(predicted_news = predict(naive_model_fake, newdata = .))
```

```{r}
fake_news%>% 
  tabyl(type, predicted_news) %>% 
  adorn_percentages("row") %>% 
  adorn_pct_formatting(digits = 2) %>%
  adorn_ns
```

```{r}
naive_model_fake
```

-   Probability(15 - word title\| article is fake) using `dnorm()`

```{r}
dnorm(15, mean = 10.42222, sd = 3.204554)
```

-   Probability(6% of words have negative associations \| article is real) using `dnorm()`

Fake - Negative

```{r}
dnorm(0.06, mean = 3.606333, sd = 1.466429)
```

Real - Negative

```{r}
dnorm(0.06, mean = 2.806556, sd = 1.190917)
```

-   Probability(no exclamation point in title \| article is real)

    -   Multiply these probabilities and save as the object **`probs_real`**

```{r}
dnorm(0, mean = 0.73333333, sd =  0.26666667)
```

```{r}
P_no_exclamation_point_given_real <-  0.73333333
P_no_exclamation_point_given_fake <- 0.97777778
```

```{r}
P_no_exclamation_given_real <- 0.97777778
P_real <- 0.6  
probs_real <- P_no_exclamation_given_real * P_real
```

```{r}
print(probs_real)
```

Lastly divide your **`probs_real`** by the sum of **`probs_real`** and **`probs_fake`** to see if you can reproduce the output from `naiveBayes()` above

```{r}

```

### Confusion Matrix - 2 points

Calculate a confusion matrix by first mutating a column to fake_news called `predicted_type` . Then, use `tabyl()` to create the matrix

```{r}
fake_news <- fake_news %>%
  mutate(predicted_type = sample(c("fake_news", "real_news"), nrow(fake_news), replace = TRUE))
```

```{r}
confusion_matrix <- fake_news %>%
  tabyl(predicted_news, predicted_type)
```

```{r}
print(confusion_matrix)
```

### How can our model be improved? - 2 points

Think about the results of the confusion matrix, is the model performing well? Try creating a new model that uses all of the features in the fake_news dataset to make a prediction on type (fake vs true). Then, create a new confusion matrix to see if the model improves.

-   I think that based off of the results above from the confusion matrix the model is not performing so well as the accuracy is quite low.
